---
title: "R Notebook"
output: html_notebook
---

## Introduction 

Goal: Build a Regression Model to predict vehicle MSRP based on various predictors. 
Dataset: Cars dataset with features including make, model, year, engine, and other properties of the car used to predict its price.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load in Data
```{r}
library(dplyr)

df = read.csv("C:/Users/caroc/OneDrive/Desktop/DS6372_Project1/data1.csv", header=TRUE, stringsAsFactors=TRUE)
head(df)
glimpse(df)
```

### Look at missing values
```{r}
colSums(is.na(df))
df[duplicated(df),]
```
- It looks like Engine.HP has 69 NAs, Engine.Cylinders has 30 NAs, Number.of.Doors has 6NAs. How do we want to deal with missing data? 
- Also looks like there are duplicate rows. Explore further to see if cam be removed. 

### Unique values for categorical variables
```{r}
library(tidyr)
df %>% select_if(is.factor) %>%
  summarise_all(~n_distinct(.)) %>% 
  pivot_longer(., everything(), names_to = "columns", values_to = "count_unique_values") %>%
  arrange(desc(count_unique_values))
```
### Explore Structure of Data Set 
```{r}
names(df)
dim(df)
str(df)
summary(df)
```
## Inspect Categorical Levels
```{r}
library(inspectdf)
library(dplyr)
df_cat <- df %>% inspect_cat()
head(df_cat)
df_cat$levels
```
- Might need to reformat some of the strings. 
- Factors as numeric? might not need to do that for these models. 


