---
title: "Applied Stat Project 1"
author: "Joey Hernandez"
date: "2023-01-22"
output: html_document
---


```{r}
library(tidyverse)
library(naniar)
dataset <- read.csv("C:/Users/Joey/Desktop/applied-stats/datasets/data1.csv")
#glimpse(dataset) # looking at the var types of each
summary(dataset) # visually looking at dataset
vis_miss(dataset) # viz looking @ missing values

```
Var. that contain NA:

Engine.HP = 69

Engine.Cylinders = 30

Number.of.Doors = 6 

Since many of the missing values seem to be found we will input them into the dataset below instead of removing those observations.

Number of Doors: All NA values here were found and input into the dataset.
```{r}
#updating NA for Doors
dataset[6931:6935,9] <- 4 # Tesla model S
dataset[4667,9] <- 2 # Ferrari FF 

```


Engine Cylinders

The Mazda RX-7, and RX-8 use a unique rotary Engine and thus technically have no cylinders. This will remain NA for now. 

The remaining NA values come from electric vehicles and have been updated to 0 to match the other electric vehicles.
```{r}
dataset[1984:1985, 6] <- 0 # chevy bolt EV
dataset[3717:3720, 6] <- 0 # volkswagen e-golf
dataset[5779:5781,6] <- 0 # mitsubishi i-miev
dataset[8374,6] <- 0 # toyota rav4-ev
dataset[c(5826,5831,5832,5834,5840,5841),6]<- 305 # chevy impala flex-fuel
dataset[c(6909,6911,6917,6919),6]<-240 # lincoln mkz

```

Engine HP

Certain cars specifically electric vehicles do not report "horsepower" but instead will report Kilowatts (KW)
so we expect to see many Missing values when dealing with Electric vehicles. 

Exceptions to this:
Chevy Spark - GM reports 140 HP
i3 - BMW reports 170-181 HP
i-MiEV - Mitsubishi reports 66 HP
e-golf - Volkswagen reports 115 HP


```{r}
dataset[8374,5] <- 'NA' # toyota RAV4-EV
dataset[8374,5]

dataset[2906:2909,5] <- 305 # lincoln Continental
dataset[4204:4207,5]<- 168 # ford escape
dataset[4915:4920,5]<- 193 # ford freestar

summary(dataset)

car_data <- dataset
```