---
always_allow_html: TRUE
output: 
  html_document: 
    code_folding: hide
  github_document: default
---

<hr>
<center><h2>Explanatory Data Analysis</h2></center>
<hr>

<h4>Loading Data set and inspecting summary statistics as well as missing data:</h4>


```{r, message=FALSE, warning=FALSE, fig.align='center'}
library(tidyverse)
library(naniar)

dataset <- read.csv("C:/Users/Joey/Desktop/applied-stats/datasets/data1.csv")

#glimpse(dataset) # looking at the var types of each

summary(dataset) # visually looking at the sum. stats of the vars in the data. 

vis_miss(dataset) # viz looking @ missing values

```


We can see there is a number of variables that contain NA's

<h4>Var. that contain NA:</h4>

- Engine.HP = 69
- Engine.Cylinders = 30
- Number.of.Doors = 6 

Since many of the missing values seem to be found using any number of automobile spec. websites we will input them into the dataset below instead of removing those observations so that we can fully utilize the features and data provided.


<h4>Number of Doors:</h4>
 All NA values here belong to Tesla Model S and Ferrari FF. We were able to find and input these values. 

```{r, message=FALSE, warning=FALSE, fig.align='center'}
#updating NA for Doors
dataset[6931:6935,9] <- 4 # Tesla model S
dataset[4667,9] <- 2 # Ferrari FF 

```



<h4>Engine Cylinders</h4>

The Mazda RX-7, and RX-8 use a unique rotary Engine and thus technically have no cylinders.

On the EV side of the data set we find the remaining NA values and they have been updated to 0 since electric vehicles do not have engine cylinders. future consideration for a substitute value would be ideal here as the automobile market gains more and more Electric vehicles. This substitute value would help to leverage a way to add accurate predictions and values using this metric when comparing many EV.

```{r, message=FALSE, warning=FALSE, fig.align='center'}

dataset[1984:1985, 6] <- 0 # chevy bolt EV
dataset[3717:3720, 6] <- 0 # volkswagen e-golf
dataset[5779:5781,6] <- 0 # mitsubishi i-miev
dataset[8374,6] <- 0 # toyota rav4-ev
dataset[8696:8715, 6] <- 0 # mazda rotary engine RX7 AND RX8



```


<h4>Engine HP</h4>

Certain cars specifically electric vehicles do not report "horsepower" but instead will report Kilowatts (KW)
so we expect to see many Missing values when dealing with Electric vehicles. 

Exceptions to this:
Chevy Spark - GM reports 140 HP
i3 - BMW reports 170-181 HP
i-MiEV - Mitsubishi reports 66 HP
e-golf - Volkswagen reports 115 HP

we could possibly create a ML algo to predict what the HP is... but since they are reported on a different scale, and are of a different type of vehicle
maybe it would be better off changing the "missing values" into 0s. 

```{r, message=FALSE, warning=FALSE, fig.align='center'}
dataset[8374,5] <- 154 # toyota RAV4-EV
dataset[2906:2909,5] <- 305 # lincoln Continental
dataset[4204:4207,5]<- 168 # ford escape
dataset[4915:4920,5]<- 193 # ford freestar
dataset[c(5826,5831,5832,5834,5840,5841),5]<- 305 # chevy impala flex-fuel
dataset[c(6909,6911,6917,6919),5]<-240 # lincoln mkz
dataset[6579, 5] <- 200 # mercedes-benz m-class
dataset[8375:8376,5]<-154 # 2013 rav4 EV and 2014
dataset[540:542, 5] <- 111 # fiat 500 e - from edmunds
dataset[9851:9855,5]<- 109 # kia soul Ev
dataset[4706:4707,5] <- 123 # honda fit Ev
dataset[c(4786,4790,4799),5] <- 143 # ford focus ev
dataset[5779,5] <- 66 # mitsubishi i-miev
dataset[6386:6395,5] <- 107 # nissan leaf 
dataset[6923,5] <- 302 # base tsla model s 2014
dataset[6922,5] <- 302 # 2nd tsla model s 2014
dataset[6925,5] <- 362 # 3rd tsla model s 2014
dataset[6924,5] <- 416 # 4th tsla model s 2014
dataset[6926:6927,5]<- 329 # tsla model s 2015
dataset[6930, 5]<- 362 # tsla model s 2015
dataset[6928,5]<- 416 # tsla model s 2015
dataset[6929,5]<- 691 # tsla model s 2015
dataset[6932,5]<-259 # tsla model s 2016
dataset[6935,5]<-315
dataset[6934:6936,5] <- 373 # tsla model s 2016
dataset[6936,5]<-373# tsla model s 2016
dataset[c(6931,6937),5]<-417# tsla model s 2016
dataset[6939,5]<- 503# tsla model s 2016
dataset[c(6933,6938),5] <- 691# tsla model s 2016

dataset[11322:11324,4] <- 'regular unleaded' # suzuki verona missing fuel type


car_data <- dataset

```


<h4>"Unknown" Data values in the Transmission.Type category</h4>
Values of "Unknown" were discovered in the data. These Unknown values are publicly and easily accessible information and we 
were able to update the database with the proper assignments.


```{r, message=FALSE, warning=FALSE, fig.align='center'}
car_data[1290:1291, 7] <- "AUTOMATIC" # oldsmobile achieva
car_data[4692,7]<- "MANUAL" # cheaper firebird
car_data[4693:4694,7]<- "AUTOMATIC" #more expensive firebird
car_data[6159,7] <- "MANUAL" # GMC Jimmy cheaper 99
car_data[6161,7] <- "AUTOMATIC" # GMC jimmy 99
car_data[6166,7] <- "MANUAL" #gmc jimmy 00
car_data[6175,7]<- "AUTOMATIC" # gmc jimmy 00
car_data[6367,7] <- "MANUAL" # chrysler le baron 
car_data[6369,7] <- "AUTOMATIC"# chrysler le baron 
car_data[c(8043,8044,8047,8048,8049,8050,8052,8054),7] <- "AUTOMATIC" # dodge ram 150

```



There are many chr and int types so we will change them into a factor where needed so that we can adequately plot the data.
```{r, message=FALSE, warning=FALSE, fig.align='center'}
glimpse(car_data)
car_data$Make <- as.factor(car_data$Make)
car_data$Engine.Fuel.Type<- as.factor(car_data$Engine.Fuel.Type) # chr
car_data$Engine.HP<- as.numeric(car_data$Engine.HP) # dbl 
car_data$Engine.Cylinders<-as.factor(car_data$Engine.Cylinders) # dbl
car_data$Transmission.Type<-as.factor(car_data$Transmission.Type) # chr
car_data$Driven_Wheels<- as.factor(car_data$Driven_Wheels) # chr
car_data$Number.of.Doors<- as.integer(car_data$Number.of.Doors) # dbl
car_data$Vehicle.Size<-as.factor(car_data$Vehicle.Size) # chr
car_data$Vehicle.Style<-as.factor(car_data$Vehicle.Style) # chr
car_data$Market.Category<-as.factor(car_data$Market.Category) # chr
```


<h4>"Not Applicable" Values:</h4>

upon investigation of the Market Category section we found many "N/A" values (3,742).
These values seem to be implicitly add N/A values, which could in fact mean there is no applicable group that it fits into. we
will change this N/A to "No Category" to avoid any confusion with Missing value data. 

```{r, message=FALSE, warning=FALSE, fig.align='center'}
# change so that we can handle the "N/A" which is NOT an NA
car_data$Market.Category<-as.character(car_data$Market.Category)
sum(car_data$Market.Category == "N/A") #3742 this is our benchmark 

# reading as not applicable we will change to no category
car_data$Market.Category[car_data$Market.Category == "N/A"] <- "No Category"

# sanity check should be 0
sum(car_data$Market.Category == "N/A")

# change to match rest of data type (factor)
car_data$Market.Category<-as.factor(car_data$Market.Category)

```
Typo Audi MPG
```{r}
view(car_data)
car_data[1120,13] <- 34

```


Finally we will remove the "model" variable since it presents us with no value.
```{r}
#glimpse(car_data) # view the df
car_data <- car_data[,-2] # rm the model column
#glimpse(car_data) # sanity check the df

```





<hr>
remaining considerations
<hr>

<h4>Duplicated Data</h4>
It would appear at first glance that there are many duplicates within this dataset, however if
you look at HP/MSRP, or other types of attributes of certain vehicles you will see subtle differences. 
While it is not an explicitly stated this is assumed to be due to various trims, and engine types/offerings that 
each vehicle offers.



feat. selection 
```{r}
library(caret)

glimpse(car_data)
first_fir <- lm(log(MSRP)~Engine.HP+Engine.Cylinders,car_data)
plot(first_fir)


model_1 <- lm(log(MSRP)~.,car_data)
plot(model_1)

set.seed(7)

fitControl <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
glmnet.fit <- train(log(MSRP)~.,
                    data = car_data,
                    method = "glmnet",
                    trControl = fitControl)

glmnet.fit
opt.pen<- glmnet.fit$finalModel$lambdaOpt
coef(glmnet.fit$finalModel,opt.pen)


car_glm<- car_data[,!names(car_data) %in% c("Engine.Cylinders6","city.mpg","Chrysler",
                                            "Ford","Hummer","Infiniti","Mazda","Volkswagen","natural",
                                            "prem_rec","front_drive","auto_manual","two_door","fourdr_SUV")]

model_2 <- lm(log(MSRP)~.,car_glm)
plot(model_2)

```



categorical variable encoding

```{r}
car_data$BMW <- ifelse(car_data$Make == "BMW",1,0) # if bmw then 1 else(not) 0
car_data$Audi <- ifelse(car_data$Make=="Audi",1,0)
car_data$FIAT <- ifelse(car_data$Make=="FIAT",1,0)
car_data$Mercedes <- ifelse(car_data$Make=="Mercedes-Benz",1,0)
car_data$Acura <- ifelse(car_data$Make=="Acura",1,0)
car_data$Alfa_Romeo <- ifelse(car_data$Make=="Alfa Romeo",1,0)
car_data$Aston_Martin <- ifelse(car_data$Make=="Aston Martin",1,0)
car_data$Bentley <- ifelse(car_data$Make=="Bentley",1,0)
car_data$Bugatti <- ifelse(car_data$Make=="Bugatti",1,0)
car_data$Buick <- ifelse(car_data$Make=="Buick",1,0)
car_data$Cadillac <- ifelse(car_data$Make=="Cadillac",1,0)
car_data$Chevrolet <- ifelse(car_data$Make=="Chevrolet",1,0)
car_data$Chrysler <- ifelse(car_data$Make=="Chrysler",1,0)
car_data$Dodge <- ifelse(car_data$Make=="Dodge",1,0)
car_data$Ferrari <- ifelse(car_data$Make=="Ferrari",1,0)
car_data$Ford <- ifelse(car_data$Make=="Ford",1,0)
car_data$Genesis <- ifelse(car_data$Make=="Genesis",1,0)
car_data$GMC <- ifelse(car_data$Make=="GMC",1,0)
car_data$Honda <- ifelse(car_data$Make=="Honda",1,0)
car_data$Land_Rover <- ifelse(car_data$Make=="Land Rover",1,0)
car_data$Lamborghini <- ifelse(car_data$Make=="Lamborghini",1,0)
car_data$Maybach <- ifelse(car_data$Make=="Maybach",1,0)
car_data$Hummer <- ifelse(car_data$Make=="Hummer",1,0)
car_data$Hyundai <- ifelse(car_data$Make=="Hyundai",1,0)
car_data$Infiniti <- ifelse(car_data$Make=="Infiniti",1,0)
car_data$Kia <- ifelse(car_data$Make=="Kia",1,0)
car_data$Mazda <- ifelse(car_data$Make=="Mazda",1,0)
car_data$Maserati <- ifelse(car_data$Make=="Maserati",1,0)
car_data$Lotus <- ifelse(car_data$Make=="Lotus",1,0)
car_data$Lincoln <- ifelse(car_data$Make=="Lincoln",1,0)
car_data$Lexus <- ifelse(car_data$Make=="Lexus",1,0)
car_data$McLaren <- ifelse(car_data$Make=="McLaren",1,0)
car_data$Mitsubishi <- ifelse(car_data$Make=="Mitsubishi",1,0)
car_data$Nissan <- ifelse(car_data$Make=="Nissan",1,0)
car_data$Oldsmobile <- ifelse(car_data$Make=="Oldsmobile",1,0)
car_data$Plymouth <- ifelse(car_data$Make=="Plymouth",1,0)
car_data$Saab <- ifelse(car_data$Make=="Saab",1,0)
car_data$Rolls_Royce <- ifelse(car_data$Make=="Rolls-Royce",1,0)
car_data$Porsche <- ifelse(car_data$Make=="Porsche",1,0)
car_data$Pontiac <- ifelse(car_data$Make=="Pontiac",1,0)
car_data$Subaru <- ifelse(car_data$Make=="Subaru",1,0)
car_data$Suzuki <- ifelse(car_data$Make=="Suzuki",1,0)
car_data$Tesla <- ifelse(car_data$Make=="Tesla",1,0)
car_data$Toyota <- ifelse(car_data$Make=="Toyota",1,0)
car_data$Volkswagen <- ifelse(car_data$Make=="Volkswagen",1,0)
car_data$Volvo <- ifelse(car_data$Make=="Volvo",1,0)
car_data$Scion <- ifelse(car_data$Make=="Scion",1,0)
car_data$Spyker <- ifelse(car_data$Make=="Spyker",1,0)

```


```{r}
# levels(car_data$Engine.Fuel.Type)
# 
 car_data$diesel <- ifelse(car_data$Engine.Fuel.Type == "diesel",1,0)
 car_data$flex_prem_rec <- ifelse(car_data$Engine.Fuel.Type=="flex-fuel (premium unleaded recommended/E85)",1,0)
 car_data$electric <- ifelse(car_data$Engine.Fuel.Type=="electric",1,0)
 car_data$flex_unleaded <- ifelse(car_data$Engine.Fuel.Type=="flex-fuel (unleaded/E85)",1,0)
 car_data$flex_prem_req <- ifelse(car_data$Engine.Fuel.Type=="flex-fuel (premium unleaded required/E85)",1,0)
 car_data$flex_fuel_natural <- ifelse(car_data$Engine.Fuel.Type=="flex-fuel (unleaded/natural gas)",1,0)
 car_data$natural <- ifelse(car_data$Engine.Fuel.Type == "natural gas",1,0)
 car_data$prem_required <- ifelse(car_data$Engine.Fuel.Type=="premium unleaded (required)",1,0)
 car_data$prem_rec <- ifelse(car_data$Engine.Fuel.Type=="premium unleaded (recommended)",1,0)
 car_data$regular_gas <- ifelse(car_data$Engine.Fuel.Type == "regular unleaded",1,0)

```

```{r}

levels(car_data$Driven_Wheels)

car_data$front_drive <- ifelse(car_data$Driven_Wheels == "front wheel drive",1,0)
car_data$rear_drive <- ifelse(car_data$Driven_Wheels =="rear wheel drive",1,0)
car_data$four_wheel <- ifelse(car_data$Driven_Wheels == "four wheel drive",1,0)
car_data$awd <- ifelse(car_data$Driven_Wheels=="all wheel drive",1,0)
dim(car_data)
```

```{r}


# levels(car_data$Transmission.Type)
car_data$direct_drive <- ifelse(car_data$Transmission.Type == "DIRECT_DRIVE", 1,0)
car_data$auto_manual <- ifelse(car_data$Transmission.Type == "AUTOMATED_MANUAL", 1,0)
car_data$automatic<- ifelse(car_data$Transmission.Type=="AUTOMATIC",1,0)
car_data$manual<- ifelse(car_data$Transmission.Type=="MANUAL",1,0)
dim(car_data)


```


```{r}


car_data$two_door <- ifelse(car_data$Number.of.Doors == "2",1,0)
car_data$four_door <- ifelse(car_data$Number.of.Doors == "4",1,0)
car_data$three_door <- ifelse(car_data$Number.of.Doors=="3",1,0)
dim(car_data)
```

```{r}




car_data$twodr_hatch <- ifelse(car_data$Vehicle.Style=="2dr Hatchback",1,0)
car_data$twodr_suv <- ifelse(car_data$Vehicle.Style=="2dr SUV",1,0)
car_data$fourdr_SUV <- ifelse(car_data$Vehicle.Style=="4dr SUV",1,0)
car_data$cargo_minivan<-ifelse(car_data$Vehicle.Style=="Cargo Minivan",1,0)
car_data$cargo_van <- ifelse(car_data$Vehicle.Style=="Cargo Van",1,0)
car_data$ext_cab_pickup <-ifelse(car_data$Vehicle.Style=="Extended Cab Pickup",1,0)
car_data$coupe<-ifelse(car_data$Vehicle.Style=="Coupe",1,0)
car_data$pass_van <-ifelse(car_data$Vehicle.Style=="Passenger Van",1,0)
car_data$crew_pickup <- ifelse(car_data$Vehicle.Style=="Crew Cab Pickup",1,0)
car_data$reg_pickup <- ifelse(car_data$Vehicle.Style=="Regular Cab Pickup",1,0)
car_data$convertible <- ifelse(car_data$Vehicle.Style=="Convertible",1,0)
car_data$suv_notop <- ifelse(car_data$Vehicle.Style=="Convertible SUV",1,0)
car_data$mini_van_pass <- ifelse(car_data$Vehicle.Style=="Passenger Minivan",1,0)
car_data$wagon <- ifelse(car_data$Vehicle.Style=="Wagon",1,0)
car_data$sedan<- ifelse(car_data$Vehicle.Style=="Sedan",1,0)
car_data$fourdr_hatch<-ifelse(car_data$Vehicle.Style=="4dr Hatchback",1,0)
#glimpse(car_data) # view the df
car_data <- car_data[,-c(1, 3, 6, 7,9,10,11)] # rm the model column
#glimpse(car_data) # sanity check the df

dim(car_data)
glimpse(car_data)
```


scaling data for use in our Regression models
```{r}
car_data$Engine.HP <- as.integer(car_data$Engine.HP)
car_data$Engine.Cylinders<-as.integer(car_data$Engine.Cylinders)
car_scaled <- car_data

car_scaled[2:8] <- as.data.frame(scale(car_scaled[2:8]))

```


viz: variables in data set
```{r, message=FALSE, warning=FALSE, fig.align='center'}



car_data%>%ggplot(aes(Market.Category))+
  geom_bar()+coord_flip()

# which make of car has the highest count? 
car_data %>% ggplot(aes(Make)) + 
  geom_bar(stat = 'count') + coord_flip()

# mpg distribution between cars
 car_data %>% ggplot(aes(city.mpg))+
   geom_histogram()
 
# hwy dist. between cars
car_data %>% ggplot(aes(highway.MPG))+
  geom_histogram() # is there a report for EV? 

# number of doors bar plot
car_data %>% ggplot(aes(Number.of.Doors))+
  geom_bar() # most cars have 4 doors (sedans)

car_data %>% ggplot(aes(Vehicle.Style))+
  geom_bar() + coord_flip() # this sedan and 4dr suv matches what we saw earlier. 

car_data %>% ggplot(aes(Transmission.Type))+
  geom_bar()  # are there any electric cars listed as something other than direct_drive? 

# Engine HP dist. 
car_data %>% ggplot(aes(Engine.HP))+
  geom_histogram()

car_data %>% ggplot(aes(Popularity))+
  geom_histogram() # not sure how this is rated.. but theres a huge part missing 
 # most tend to be found in 0 - just above 2000

# representation of drive type (AWD,FWD, 4WD etc)
car_data %>% ggplot(aes(Driven_Wheels))+
  geom_bar()

# right skewed data for Year (which makes sense since we expect to be selling "newer" vehicles)
car_data %>% ggplot(aes(Year)) +
  geom_bar()

car_data %>% ggplot(aes(Engine.Fuel.Type))+
  geom_bar() + coord_flip()

```

this chunk is to look at MSRP broken into different ranges and finally logged
```{r, message=FALSE, warning=FALSE, fig.align='center'}
car_data %>% ggplot(aes(MSRP))+
  geom_histogram()

car_data %>% filter(MSRP <=10000) %>% ggplot(aes(MSRP)) +
  geom_histogram()

car_data %>% filter(MSRP <=20000) %>% ggplot(aes(MSRP)) +
  geom_histogram()

car_data %>% filter(MSRP <=40000) %>% ggplot(aes(MSRP)) +
  geom_histogram()

car_data %>% filter(MSRP <=60000) %>% ggplot(aes(MSRP)) +
  geom_histogram()

car_data %>% filter(MSRP <=80000) %>% ggplot(aes(MSRP)) +
  geom_histogram()

car_data %>% filter(MSRP <=100000) %>% ggplot(aes(MSRP)) +
  geom_histogram()

car_data %>% ggplot(aes(log(MSRP)))+
  geom_histogram()

max(car_data$MSRP)
```


viz - of some intuitive relationships
should be noted that since we are dealing with MSRP we may need to scale the data so that it can be visualized properly.
Numerical data
```{r, message=FALSE, warning=FALSE, fig.align='center'}
glimpse(car_data)
# numerical 
par(mfrow = c(2,2))
# city mpg
plot(car_data$city.mpg,car_data$MSRP, main = "city mpg")
#hwy mpg
plot(car_data$highway.MPG, car_data$MSRP, main = "hwy mpg")
# popularity
plot(car_data$Popularity, car_data$MSRP, main = "popular") # this looks irrelevant ?

plot(car_data$Engine.HP, car_data$MSRP, main="Engine HP")

```



```{r, message=FALSE, warning=FALSE, fig.align='center'}
car_data %>% filter(Year >= 2000) %>% ggplot(aes(Year, MSRP))+
  geom_point()

car_data %>% filter(Year <= 2000 ) %>% ggplot(aes(Year, MSRP))+
  geom_point()

```


categorical data -1 
```{r, message=FALSE, warning=FALSE, fig.align='center'}
library(ggplot2)

# categorical 
par(mfrow = c(2,3))
# Veh. Style
plot(car_data$Vehicle.Style, log(car_data$MSRP), main = "vehicle style")
# make
plot(car_data$Make,log(car_data$MSRP), main = "make")
# year 
plot(car_data$Year, log(car_data$MSRP), main = "year")

# fuel type
plot(car_data$Engine.Fuel.Type, log(car_data$MSRP), main = "fuel type")
# HP 
plot(car_data$Engine.HP, log(car_data$MSRP), main = "horse power")
# cylinders
plot(car_data$Engine.Cylinders, log(car_data$MSRP), main = "cylinders")
```

categorical data -2
```{r, message=FALSE, warning=FALSE, fig.align='center'}
par(mfrow = c(2,3))
# transmission type
plot(car_data$Transmission.Type, log(car_data$MSRP), main = "trans. type")
# wheel drive
plot(car_data$Driven_Wheels, log(car_data$MSRP), main = "wheel drive")
# number of doors
plot(car_data$Number.of.Doors, log(car_data$MSRP), main = "num of doors")
# market category
plot(car_data$Market.Category, log(car_data$MSRP), main = "market category")  #there is big chunk of missing values this may be unreliable
# size
plot(car_data$Vehicle.Size, log(car_data$MSRP), main = "vehicle size")
# style
plot(car_data$Vehicle.Style, log(car_data$MSRP), main = "vehicle style")

```




Feature Selection scaled data - looks like trash
```{r}
library(caret)
glimpse(car_scaled)

set.seed(7)

fitControl <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
glmnet.fit <- train(MSRP~.,
                    data = car_scaled,
                    method = "glmnet",
                    trControl = fitControl)

glmnet.fit
opt.pen<- glmnet.fit$finalModel$lambdaOpt
coef(glmnet.fit$finalModel,opt.pen)

car_glm_scaled <- car_scaled[,!names(car_scaled) %in% c("Ford", "Hummer","Saab","Pontiac","Volkswagen", "Volvo", "Scion","regular_gas", "automatic", "four_door","fourdr_SUV")]

model_fit <- lm(MSRP ~. ,car_glm_scaled)
summary(model_fit)
plot(model_fit)

```